---
title: "Data 621 - Final Project Proposal"
author: "Dhairav Chhatbar, Mael Illien, Salma Elshahawy"
date: "11/17/2020"
output: 
  html_document:
    code_folding: hide
    theme: cosmo
    highlight: tango
    toc: true
    number_section: false
    toc_float:
      collapsed: true
      smooth_scroll: true
    df_print: paged
---

```{r message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(skimr)
library(readr)
```

## Dataset

For this project, we propose to explore the Default of Credit Card Clients Data Set from the UCI Machine Learning Repository hosted on [Kaggle](https://www.kaggle.com/uciml/default-of-credit-card-clients-dataset?select=UCI_Credit_Card.csv). This dataset contains information on default payments, demographic factors, credit data, history of payment, and bill statements of credit card clients in Taiwan from April 2005 to September 2005.

The dataset has around ~30,000 observations and 25 columns. The variables are as the following:

```{r message=FALSE, warning=FALSE}
dataset = read_csv('https://raw.githubusercontent.com/salma71/Data_621/master/Project_Proposal/UCI_Credit_Card.csv')
skim(dataset)
```

The variables discription are as follows:

| ID                         	| ID of each client                                                                                                                                                                               	|
|----------------------------	|-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------	|
| LIMIT_BAL                  	| Amount of given credit in NT dollars (includes individual and family/supplementary credit                                                                                                       	|
| SEX                        	| Gender (1=male, 2=female)                                                                                                                                                                       	|
| EDUCATION                  	| (1=graduate school, 2=university, 3=high school, 4=others, 5=unknown, 6=unknown)                                                                                                                	|
| MARRIAGE                   	| Marital status (1=married, 2=single, 3=others)                                                                                                                                                  	|
| AGE                        	| Age in years                                                                                                                                                                                    	|
| PAY_0                      	| Repayment status in September, 2005 (-1=pay duly, 1=payment delay for one month, 2=payment delay for two months, … 8=payment delay for eight months, 9=payment delay for nine months and above) 	|
| PAY_2                      	| Repayment status in August, 2005 (scale same as above)                                                                                                                                          	|
| PAY_3                      	| Repayment status in July, 2005 (scale same as above)                                                                                                                                            	|
| PAY_4                      	| Repayment status in June, 2005 (scale same as above)                                                                                                                                            	|
| PAY_5                      	| Repayment status in May, 2005 (scale same as above)                                                                                                                                             	|
| PAY_6                      	| Repayment status in April, 2005 (scale same as above)                                                                                                                                           	|
| BILL_AMT1                  	| Amount of bill statement in September, 2005 (NT dollar)                                                                                                                                         	|
| BILL_AMT2                  	| Amount of bill statement in August, 2005 (NT dollar)                                                                                                                                            	|
| BILL_AMT3                  	| Amount of bill statement in July, 2005 (NT dollar)                                                                                                                                              	|
| BILL_AMT4                  	| Amount of bill statement in June, 2005 (NT dollar)                                                                                                                                              	|
| BILL_AMT5                  	| Amount of bill statement in May, 2005 (NT dollar)                                                                                                                                               	|
| BILL_AMT6                  	| Amount of bill statement in April, 2005 (NT dollar)                                                                                                                                             	|
| PAY_AMT1                   	| Amount of previous payment in September, 2005 (NT dollar)                                                                                                                                       	|
| PAY_AMT2                   	| Amount of previous payment in August, 2005 (NT dollar)                                                                                                                                          	|
| PAY_AMT3                   	| Amount of previous payment in July, 2005 (NT dollar)                                                                                                                                            	|
| PAY_AMT4                   	| Amount of previous payment in June, 2005 (NT dollar)                                                                                                                                            	|
| PAY_AMT5                   	| Amount of previous payment in May, 2005 (NT dollar)                                                                                                                                             	|
| PAY_AMT6                   	| Amount of previous payment in April, 2005 (NT dollar)                                                                                                                                           	|
| default.payment.next.month 	| Default payment (1=yes, 0=no)                                                                                                                                                                   	|
|

As shown from the peek into the dataset, mostt of the variables are severly skewed, have low standard deviation (Low variance) and relatively high mean(high bias) this in return would result in underfitting model and less generalization. In this case we would use some of the resampling methods like, Bootstrap aggtigation method (Bagging) to combat the underfitting (Bias-variance tradeoff). 

## Abstract

 - The goal of this project is to estimate the propability of default payment by credit card history. 
 - Use the advanced techniques of classification to predict if a client has high probability for a default. 
 - To investigate if there is a trend for a default between the time that the dataset was collected.
 
## Introduction

For this dataset, we are intending to develop a statistical model for probabilty of default, how many clients already been defaulted. Additionally, we would create exhaustive visualizations between dependent variable and the independent variables ,where we can infer the correlations between variables. In addition investigate the variables distributions to select the best classifer. 

Based on the previous step, we would start our data preprocessing and cleaning to prepare the data for modeling. We would approach the least distructive preprocessing methods to preserve variables distributions to make the model as reliable as we can. 

After that, we would extend our work by building a several classifers using different feature selection techniques then derive the performance metrics for each model. Finally, we would select model with the best performance and create predictions based on it. 


## Acknowledgements

Lichman, M. (2013). (UCI Machine Learning Repository)[http://archive.ics.uci.edu/ml]. Irvine, CA: University of California, School of Information and Computer Science.

The original dataset can be found here at the (UCI Machine Learning Repository)[https://archive.ics.uci.edu/ml/datasets/default+of+credit+card+clients].



