---
title: "Untitled"
author: "Salma Elshahawy"
date: "12/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r}
library(xts)
library(DMwR2)
library(quantmod)
library(TTR)
library(performanceEstimation) 
library(nnet)
library(e1071)
library(kernlab)
library(earth)
library(randomForest)
library(ggplot2)
library(dplyr)
library(plotly)
library(hrbrthemes)
```

```{r}
library(readr)
dataset <- read_csv('https://raw.githubusercontent.com/salma71/Data_621/master/Project_Proposal/stocks_combined.csv')
tickers <- read_csv('https://raw.githubusercontent.com/salma71/Data_621/master/Project_Proposal/tickers.csv')
```
```{r}
head(tickers)
```


```{r}
head(dataset)
```

```{r}
dataset$date <- as.Date(anytime::anydate(dataset$date))
head(dataset)
```

```{r}
# library(dygraphs)
# 
# # Usual area chart
# don <- xts( x=dataset$close, order.by=dataset$date)
# 
# # Turn it interactive with ggplotly
# p <- dygraph(don)
# p
```

It seems that the closing price tends to increase by time. Let's investigate closing price by ticker.

```{r}
# Get quantmod
if (!require("quantmod")) {
    install.packages("quantmod")
    library(quantmod)
}

```

```{r}
library(timetk)
library(dplyr)
# data %>%
#   dplyr::group_by(ticker) %>%
#   plot_time_series(date, close, 
#                    .facet_ncol = 2,
#                    .facet_scale = 'free',
#                    .interactive = TRUE)
```

```{r}
stocks = tickers$Ticker
stockEnv <- new.env()
symbols <- getSymbols(stocks, src='yahoo', env=stockEnv)
for (stock in ls(stockEnv)){
    chartSeries(stockEnv[[stock]], theme="white", name=stock,
        TA="addVo();addBBands();addCCI();addSMA(20, col='blue');
        addSMA(5, col='red');addSMA(50, col='black')", subset='2014-07-01 /2019-01-09') 
}
```




```{r}
ggplot(dataset, aes(x=date, y=close, col=ticker)) + 
  geom_line() + 
  theme_bw()

```

Filter out the tickers to get the ones that has significant trends form mid 2016 to 2019

```{r}
dataset %>%
  subset(date > "2016-06-03") %>%
  filter(close > 150) %>%
  ggplot(aes(x=date, y=close, col=ticker)) + 
  geom_line() + 
  theme_bw()
```


It seems that the major tickers that has significant trends over the years `BA`, `UNH` increasing trends. On the otherhand, `V`, `GS` has a decreasing trend in the closing price. 

I will investigate each one four main tickers. 

```{r}
stocks <- dataset %>% 
  filter(ticker %in% c("BA", "UNH","MDFT", "GS", "AAPL"))
head(stocks)
```



```{r}
library(plotly)
# cutom colors
i <- list(line = list(color = '#17BECF'))
d <- list(line = list(color = '#7F7F7F'))

fig <- stocks %>%
  filter(ticker == 'AAPL' & date > "2016-06-03") %>%
  plot_ly(x = ~date, type = 'candlestick',
          open = ~open, close = ~close,
          high = ~high, low = ~low,
          increasing = i, decreasing = d) 
fig <- fig %>%
  layout(title = "Apple Stock",
         yaxis = list(title = 'closing price'),
         xaxis = list(rangeslider = list(visible = F)))

fig
```

```{r}
fig <- stocks %>%
  filter(ticker == 'UNH' & date > "2016-06-03") %>%
  plot_ly(x = ~date, type = 'candlestick',
          open = ~open, close = ~close,
          high = ~high, low = ~low,
          increasing = i, decreasing = d) 
fig <- fig %>%
  layout(title = "UNH Stock",
         yaxis = list(title = 'closing price'),
         xaxis = list(rangeslider = list(visible = F)))

fig
```


```{r}
getSymbols(Symbols = c("BA", "UNH","MSFT", "GS", "AAPL"), env = stockEnv, src = "yahoo")
```

```{r}
temp <- as.xts(data.frame(BA = BA[, "BA.Close"], UNH = UNH[, "UNH.Close"], GS = GS[, "GS.Close"], AAPL = AAPL[, "AAPL.Close"]))
head(temp)
```

```{r}
plot(as.zoo(temp), screens = 1, lty = 1:3, xlab = "Date", ylab = "Price")
legend("topleft", c("BA", "UNH","GS", "AAPL"), lty = 1:3, cex = 0.5)
```

```{r}
stock_return = apply(temp, 1, function(x) {x / temp[1,]}) %>% 
                                    t %>% as.xts

head(stock_return)
```


```{r}
plot(as.zoo(stock_return), screens = 1, lty = 1:3, xlab = "Date", ylab = "Return")
legend("topleft", c("BA", "UNH","GS", "AAPL"), lty = 1:3, cex = 0.5)
```



