---
title: "Data 621 - Assignment 1"
author: "Dhairav Chhatbar"
date: "9/9/2020"
output: html_document
---

```{r, message=FALSE, warning=FALSE}
library(skimr)
library(ggplot2)
library(ggcorrplot)
library(dplyr)
library(tidyr)
library(PerformanceAnalytics)
library(DescTools)
library(cowplot)
```


```{r}
data_test <- read.csv("https://raw.githubusercontent.com/salma71/Data_621/master/HW_1/datasets/moneyball-evaluation-data.csv", header = TRUE)
data_train <- read.csv("https://raw.githubusercontent.com/salma71/Data_621/master/HW_1/datasets/moneyball-training-data.csv", header = TRUE)
```

## Data Exploration
1. Impute missing values
2. Combine/new variables
3. Transforms

Outliers & Leverage Points

In summary, an outlier is a point whose standardized residual falls outside the interval from –2 to 2. Recall that a bad leverage point is a leverage point which is also an outlier. Thus, a bad leverage point is a leverage point whose standar- dized residual falls outside the interval from –2 to 2. On the other hand, a good leverage point is a leverage point whose standardized residual falls inside the interval from –2 to 2.

Recall that the rule for simple linear regression for classifying a point as a leverage point is hii > 4/n . 

## Build Models
## Select Models

```{r}
dim(data_train)
```

```{r}
# list types for each attribute
sapply(data_train, class)
```

```{r}
skim(data_train)
```

```{r message=FALSE, warning=FALSE}
library(kableExtra)
data_train %>% 
  gather(variable, value) %>%
  filter(is.na(value)) %>%
  group_by(variable) %>%
  tally() %>%
  mutate(percent = n / nrow(data_train) * 100) %>%
  mutate(percent = paste0(round(percent, ifelse(percent < 10, 1, 0)), "%")) %>%
  arrange(desc(n)) %>%
  rename(`Variable Missing Data` = variable,
         `Number of Records` = n,
         `Share of Total` = percent) %>%
  kable() %>%
  kable_styling()
```

The hit by pitcher variable is missing over 90% of it’s data. We will exclude it from consideration in our model.

Caught stealling a base (TEAM_BASERUN_CS) is next on the list. It may be possible to predict it using TEAM_BASERUN_SB since they are strongly correlated, but there are 131 times they both are missing data.

MI: TRY
- Drop variable if unlikely event in real life (HBP: hit by pitcher and CS: caught stealing base are rare)
- Impute (try: mean, 0, drop)


```{r fig.height=8, fig.width=10, message=FALSE, warning=FALSE}
#missing_per <- as.data.frame(t(data_train %>% select(everything()) %>% #summarise_all(funs(sum(is.na(.))))/nrow(data_train))) %>% arrange(desc(V1))
#missing_per
data_train %>% 
  select(-INDEX) %>% 
  gather() %>% 
  ggplot( aes(x= value)) + 
  geom_density(fill='pink') + 
  facet_wrap(~key, scales = 'free')
```

Let's take a closer look at the `TEAM_BASERUN_SB` 

```{r fig.height=8, fig.width=10, message=FALSE, warning=FALSE}
data_train %>% 
  ggplot(aes(TEAM_BASERUN_SB)) + 
  geom_histogram(bins = 50, fill = 'pink') +
  geom_vline(aes(xintercept = mean(TEAM_BASERUN_SB, na.rm = T)), col = "red", lty = 2) +
  geom_vline(aes(xintercept = median(TEAM_BASERUN_SB, na.rm = T)), col = "blue", lty = 2) +
  labs(x = element_blank(),
       y = "Count",
       title = "Distribution of Stolen Bases",
       caption = "* Red line is the mean value and blue is the median") + 
  theme_classic()
```

```{r fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
ggplot(data_train, aes(x=TARGET_WINS)) + 
    geom_histogram(aes(y=..density..),binwidth = 3, colour="black", fill="red") +
    geom_density(alpha=.8, fill="pink") + 
  theme_classic()
```

```{r fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
ggplot(data_train, aes(x=INDEX, y=TARGET_WINS)) + 
    geom_point(colour="black") +
    theme_classic()
```


```{r fig.height=8, fig.width=10, message=FALSE, warning=FALSE}
to_plot <- data_train %>% 
  dplyr::select(-INDEX)
q <- cor(to_plot)
ggcorrplot(q, type = "upper", outline.color = "white",
           ggtheme = theme_classic,
           colors = c("#6D9EC1", "white", "#E46726"),
           lab = TRUE, show.legend = FALSE, tl.cex = 8, lab_size = 3) 
```

Correlations with Response Variable
Let’s take a look at how the predictors are correlated with the response variable:
```{r fig.height=8, fig.width=10, message=FALSE, warning=FALSE}
data_train %>%
  gather(variable, value, -TARGET_WINS) %>%
  ggplot(., aes(value, TARGET_WINS)) + 
  geom_point(fill = "pink", color="pink") + 
  geom_smooth(method = "lm", se = FALSE, color = "red") + 
  facet_wrap(~variable, scales ="free", ncol = 4) +
  labs(x = element_blank(), y = "Wins")
```

It looks like the `TEAM_PITCHING_BB`, `TEAM_BATTING_H` , `TEAM_BATTING_BB`, `TEAM_BATTING_2B`and `TEAM_BATTING_HR` have the strongest correlations with wins. None of these correlations are particularly strong. This suggests there is a lot of ‘noise’ in these relationships due to outliers.

MI: 
- TEAM_PITCHING_HR (homeruns pitched) is slightly correlated win the number of wins. This is clearly unexpected because pitched homeruns are a negative for the pitching team. 
- TEAM_PITCHING_SO (strikeouts by pitcher) likely suffers from outliers because we'd expect strikeouts to be a positive thing for the pitching team. There is a large influential point.
- TEAM_PITCHING_BB (walks allowed by pitcher) might suffer from outliers. We'd expected a negative correlation since a walk allowed is a negative for the pitching team.


### Data Prep

MI:
- New variables: TEAM_BATTING_1B: TEAM_BATTING_H - SUM(TEAM_BATTING_i) where i = 2B,3B,HR
- Transforms

```{r}
# New variable: TEAM_BATTING_1B
base_data <- select(data_train, -INDEX)
mod_data <- base_data %>% mutate(TEAM_BATTING_1B = base_data$TEAM_BATTING_H - select(., TEAM_BATTING_2B:TEAM_BATTING_HR) %>% rowSums(na.rm = FALSE))
mod_data
```
```{r, warning=FALSE, message=FALSE, fig.height=15, fig.width=5}
# Data Transformations

data_transformed <- mod_data

#Log transform TEAM_BASERUN_CS
data_transformed$TEAM_BASERUN_CS_tform <-log(data_transformed$TEAM_BASERUN_CS)

baserun_cs <- ggplot(data_transformed, aes(x=TEAM_BASERUN_CS)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="red") +
    geom_density(alpha=.8, fill="pink") + 
  theme_classic() + labs(title = "TEAM_BASERUN_CS")

baserun_cs_tf <- ggplot(data_transformed, aes(x=TEAM_BASERUN_CS_tform)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="red") +
    geom_density(alpha=.8, fill="pink") + 
  theme_classic() + labs(title = "TEAM_BASERUN_CS Log Transformed")

#Log transform TEAM_BASERUN_SB
data_transformed$TEAM_BASERUN_SB_tform <-log(data_transformed$TEAM_BASERUN_SB)

baserun_sb <- ggplot(data_transformed, aes(x=TEAM_BASERUN_SB)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="red") +
    geom_density(alpha=.8, fill="pink") + 
  theme_classic() + labs(title = "TEAM_BASERUN_SB")

baserun_sb_tf <- ggplot(data_transformed, aes(x=TEAM_BASERUN_SB_tform)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="red") +
    geom_density(alpha=.8, fill="pink") + 
  theme_classic() + labs(title = "TEAM_BASERUN_SB Log Transformed")


#Log transform TEAM_BATTING_3B
data_transformed$TEAM_BATTING_3B_tform <-log(data_transformed$TEAM_BATTING_3B)

batting_3b <- ggplot(data_transformed, aes(x=TEAM_BATTING_3B)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="red") +
    geom_density(alpha=.8, fill="pink") + 
  theme_classic() + labs(title = "TEAM_BATTING_3B")

batting_3b_tf <- ggplot(data_transformed, aes(x=TEAM_BATTING_3B_tform)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="red") +
    geom_density(alpha=.8, fill="pink") + 
  theme_classic() + labs(title = "TEAM_BATTING_3B Log Transformed")

#BoxCoxtransform TEAM_BATTING_BB
data_transformed$TEAM_BATTING_BB_tform <- BoxCox(data_transformed$TEAM_BATTING_BB, BoxCoxLambda(data_transformed$TEAM_BATTING_BB))

batting_bb <- ggplot(data_transformed, aes(x=TEAM_BATTING_BB)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="red") +
    geom_density(alpha=.8, fill="pink") + 
  theme_classic() + labs(title = "TEAM_BATTING_BB")

batting_bb_tf <- ggplot(data_transformed, aes(x=TEAM_BATTING_BB_tform)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="red") +
    geom_density(alpha=.8, fill="pink") + 
  theme_classic() + labs(title = "TEAM_BATTING_BB BoxCox Transformed")


#BoxCoxtransform TEAM_BATTING_H
data_transformed$TEAM_BATTING_H_tform <- BoxCox(data_transformed$TEAM_BATTING_H, BoxCoxLambda(data_transformed$TEAM_BATTING_H))

batting_h <- ggplot(data_transformed, aes(x=TEAM_BATTING_H)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="red") +
    geom_density(alpha=.8, fill="pink") + 
  theme_classic() + labs(title = "TEAM_BATTING_H")

batting_h_tf <- ggplot(data_transformed, aes(x=TEAM_BATTING_H_tform)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="red") +
    geom_density(alpha=.8, fill="pink") + 
  theme_classic() + labs(title = "TEAM_BATTING_H BoxCox Transformed")

#BoxCoxtransform TEAM_BATTING_1B
data_transformed$TEAM_BATTING_1B_tform <- BoxCox(data_transformed$TEAM_BATTING_1B, BoxCoxLambda(data_transformed$TEAM_BATTING_1B))

batting_1b <- ggplot(data_transformed, aes(x=TEAM_BATTING_1B)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="red") +
    geom_density(alpha=.8, fill="pink") + 
  theme_classic() + labs(title = "TEAM_BATTING_1B")

batting_1b_tf <- ggplot(data_transformed, aes(x=TEAM_BATTING_1B_tform)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="red") +
    geom_density(alpha=.8, fill="pink") + 
  theme_classic() + labs(title = "TEAM_BATTING_1B BoxCox Transformed")


#BoxCoxtransform TEAM_FIELDING_E
data_transformed$TEAM_FIELDING_E_tform <- BoxCox(data_transformed$TEAM_FIELDING_E, BoxCoxLambda(data_transformed$TEAM_FIELDING_E))

fielding_e <- ggplot(data_transformed, aes(x=TEAM_FIELDING_E)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="red") +
    geom_density(alpha=.8, fill="pink") + 
  theme_classic() + labs(title = "TEAM_FIELDING_E")

fielding_e_tf <- ggplot(data_transformed, aes(x=TEAM_FIELDING_E_tform)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="red") +
    geom_density(alpha=.8, fill="pink") + 
  theme_classic() + labs(title = "TEAM_FIELDING_E BoxCox Transformed")


#Log transform TEAM_PITCHING_BB
data_transformed$TEAM_PITCHING_BB_tform <-log(data_transformed$TEAM_PITCHING_BB)

pitching_bb <- ggplot(data_transformed, aes(x=TEAM_PITCHING_BB)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="red") +
    geom_density(alpha=.8, fill="pink") + 
  theme_classic() + labs(title = "TEAM_PITCHING_BB")

pitching_bb_tf <- ggplot(data_transformed, aes(x=TEAM_PITCHING_BB_tform)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="red") +
    geom_density(alpha=.8, fill="pink") + 
  theme_classic() + labs(title = "TEAM_PITCHING_BB Log Transformed")


#BoxCoxtransform TEAM_PITCHING_H
data_transformed$TEAM_PITCHING_H_tform <- BoxCox(data_transformed$TEAM_PITCHING_H, BoxCoxLambda(data_transformed$TEAM_PITCHING_H))

pitching_h <- ggplot(data_transformed, aes(x=TEAM_PITCHING_H)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="red") +
    geom_density(alpha=.8, fill="pink") + 
  theme_classic() + labs(title = "TEAM_PITCHING_H")

pitching_h_tf <- ggplot(data_transformed, aes(x=TEAM_PITCHING_H_tform)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="red") +
    geom_density(alpha=.8, fill="pink") + 
  theme_classic() + labs(title = "TEAM_PITCHING_H BoxCox Transformed")


#Log transform TEAM_PITCHING_SO
data_transformed$TEAM_PITCHING_SO_tform <-log(data_transformed$TEAM_PITCHING_SO)

pitching_so <- ggplot(data_transformed, aes(x=TEAM_PITCHING_SO)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="red") +
    geom_density(alpha=.8, fill="pink") + 
  theme_classic() + labs(title = "TEAM_PITCHING_SO")

pitching_so_tf <- ggplot(data_transformed, aes(x=TEAM_PITCHING_SO_tform)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="red") +
    geom_density(alpha=.8, fill="pink") + 
  theme_classic() + labs(title = "TEAM_PITCHING_SO")

plot_grid(baserun_cs, baserun_cs_tf, baserun_sb, baserun_sb_tf,
          batting_3b, batting_3b_tf, batting_bb, batting_bb_tf,
          batting_h, batting_h_tf, batting_1b, batting_1b_tf, 
          fielding_e, fielding_e_tf, pitching_bb, pitching_bb_tf, 
          pitching_h, pitching_h_tf, pitching_so, pitching_so_tf, 
          ncol = 2)

```


### Build Models


```{r}
#data with all original variables
m1_data <- mod_data %>% select(-TEAM_BATTING_1B)

m1 <- lm(TARGET_WINS ~., data = m1_data)
summary(m1)
plot(m1)

#Issue: only 2 significant variables (fielding variables)
```


```{r}
#data without TEAM_BATTING_HBP, TEAM_BASERUN_CS, and TEAM_FIELDING_DP
m2_data <- mod_data %>% select(-TEAM_BATTING_HBP, -TEAM_BASERUN_CS, -TEAM_FIELDING_DP)

m2 <- lm(TARGET_WINS ~., data = m2_data)
summary(m2)
plot(m2)



```

```{r}
#data without TEAM_BATTING_HBP, TEAM_BASERUN_CS, and TEAM_FIELDING_DP
m3_data <- mod_data %>% select(-TEAM_BATTING_HBP, -TEAM_BASERUN_CS, -TEAM_FIELDING_DP, -TEAM_BATTING_H)

m3 <- lm(TARGET_WINS ~., data = m3_data)
summary(m3)
plot(m3)


```


```{r}
#data with all transformed variables
m4_data <- select(data_transformed,-TEAM_BASERUN_CS, -TEAM_BASERUN_SB, -TEAM_BATTING_3B, -TEAM_BATTING_BB, -TEAM_BATTING_H, -TEAM_FIELDING_E, -TEAM_PITCHING_BB, -TEAM_PITCHING_H, -TEAM_PITCHING_SO, -TEAM_BATTING_HBP, -TEAM_BASERUN_CS, -TEAM_BATTING_1B)

m4 <- lm(TARGET_WINS ~., data = m4_data)
summary(m4)
plot(m4)


```










